Covariate analysis of ROSMAP RNASeq

### 2016-01-22 code/Rmd/ROSMAP_RNASeq.Rmd

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Remove samples with no cogdx, RIN and PMI scores
METADATA <- METADATA %>%
  mutate(Sampleid = paste(Sampleid, Batch, sep = '_')) %>%
  filter(Sampleid %in% colnames(COUNT)) %>%
  filter(!is.na(cogdx)) %>%
  filter(!is.na(RINcontinuous)) %>%
  filter(!is.na(pmi))  
```

```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$Sampleid, colnames(EXPR))

@@

@@

```{r addVarsBack}
varsToAddBackIn <- c("cogdx1",  "cogdx2", "cogdx3", "cogdx4", "cogdx5", "cogdx6")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + (FIT$coefficients[,varsToAddBackIn] %*% t(data.matrix(DESIGN[,varsToAddBackIn])))
RESIDUAL.GENE_EXPRESSION = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsemblGeneID') %>%
  tidyr::separate(col = EnsemblGeneID, into = c('ensembl_gene_id','position'), sep = '\\.')
```

>>>>

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Remove samples with no cogdx, RIN and PMI scores
METADATA <- METADATA %>%
  mutate(Sampleid = paste(Sampleid, Batch, sep = '_')) %>%
  filter(Sampleid %in% colnames(COUNT)) %>%
  filter(!is.na(cogdx)) %>%
  filter(!is.na(RINcontinuous)) %>%
  filter(!is.na(pmi)) %>%
  filter(!(Sampleid %in% c('380_120503_1')))
```
Sample id 380\_120503\_1 is considered as outlier and removed
```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$Sampleid, colnames(COUNT))

@@

### Outliers detection
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(EXPR, scale.=T)

# Plot first 2 PCs
plotdata <- data.frame(Sampleid=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, METADATA)

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Batch), shape=as.factor(cogdx), size=as.numeric(age_death)))
p <- p + theme_bw() + theme(legend.position="top") 
p <- p + geom_text(aes(label= Sampleid), size=4, hjust=1)
p

COVARIATES.tmp = data.matrix(COVARIATES)
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(EXPR)))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE, 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```

@@

```{r addVarsBack}
varsToAddBackIn <- c("cogdx1",  "cogdx2", "cogdx3", "cogdx4", "cogdx5", "cogdx6")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + t(data.matrix(DESIGN[,varsToAddBackIn]) %*% t(FIT$coefficients[,varsToAddBackIn]))
RESIDUAL.GENE_EXPRESSION = rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, 'EnsemblGeneID') %>%
  tidyr::separate(col = EnsemblGeneID, into = c('ensembl_gene_id','position'), sep = '\\.')
```
